---
title: "Numeric EDA"
output: github_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load libraries}
library(tidyverse)
```

## Let's import data

```{r}
weather_df =  
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>% #CREATE NEW MONTH VAR
  select(name, id, everything())
```

## `group_by()`

Doesn't change data set at all, but displays number of distinct groups of specified variable(s). Can group by multiple variables.

```{r}
weather_df %>%
  group_by(name, month)
```
In this case, this returns [36] - 12 months for 3 groups.

## `summarize()`

Let's group and then summarize!

```{r}
weather_df %>%
  group_by(name) %>%
  summarize(
    n_obs = n() #n_obs is name of new variable, n() returns counts for each group
  )

weather_df %>%
  group_by(name, month) %>%
  summarize(
    n_obs = n() #n_obs is name of new variable, n() returns counts for each group
  )
```

Can we count in other ways? Yes, using `count()`. DON'T use `table()` for counts.

```{r}
weather_df %>%
  count(name)

weather_df %>%
  count(name, month)
```

You can have multiple summaries:

```{r}
weather_df %>%
  group_by(month) %>%
  summarize(
    n_obs = n(),
    n_dist = n_distinct(date)
  )
```

## 2x2 tables

```{r}
weather_df %>% 
  mutate(
    cold = case_when(
      tmax <  5 ~ "cold",
      tmax >= 5 ~ "not_cold",
      TRUE      ~ ""
  )) %>% 
  filter(name != "Waikiki_HA") %>% 
  group_by(name, cold) %>% 
  summarize(count = n())
```

This is a “tidy” table, and it’s also a data frame. You could re-organize into a more standard (non-tidy) 2x2 table using pivot_wider, or you could use janitor::tabyl (BEST):

```{r}
#with pivot_wider
weather_df %>% 
  mutate(cold = case_when(
    tmax <  5 ~ "cold",
    tmax >= 5 ~ "not_cold",
    TRUE     ~ ""
  )) %>% 
  filter(name != "Waikiki_HA") %>% 
  group_by(name, cold) %>% 
  summarize(
    n_obs = n()) %>%
  pivot_wider(
    names_from = cold,
    values_from = n_obs)

#with janitor::tabyl (QUICKER + EASIER METHOD)
weather_df %>% 
  mutate(cold = case_when(
    tmax <  5 ~ "cold",
    tmax >= 5 ~ "not_cold",
    TRUE     ~ ""
  )) %>% 
  filter(name != "Waikiki_HA") %>% 
  janitor::tabyl(name, cold) #janitor package has tons of functions
```






