---
title: "Visualization"
subtitle: "Part 2"
output: github_document
---

```{r setup}
library(tidyverse)
library(ggridges)
```

## Load the weather data

We’ll be working with NOAA weather data, which is downloaded using rnoaa::meteo_pull_monitors function in the code chunk below; similar code underlies the weather dataset used elsewhere in the course. Because this process can take some time, I’ll cache the code chunk:

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

## Scatterplot

But better this time!

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs( #label graphics
    x = "Minimum Daily Temp (C)",
    y = "Maximum Daily Temp (C)",
    title = "Scatterplot of daily temp extremes",
    caption = "Data come from the rnoaa package"
  ) +
  scale_x_continuous(
    breaks = c(-10, 0, 15),
    labels = c("-10C", "0", "15")
  ) +
  scale_y_continuous(
    trans = "sqrt"
  )
```

Playing with color (don't do this unless you need to):
```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs( #label graphics
    x = "Minimum Daily Temp (C)",
    y = "Maximum Daily Temp (C)",
    title = "Scatterplot of daily temp extremes",
    caption = "Data come from the rnoaa package"
  ) +
  scale_color_hue(
    name = "Location", #rename graphic legend
    h = c(100, 300)) #designate colors
```

Playing with color (USE **VIRIDIS** METHOD INSTEAD):
```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs( #label graphics
    x = "Minimum Daily Temp (C)",
    y = "Maximum Daily Temp (C)",
    title = "Scatterplot of daily temp extremes",
    caption = "Data come from the rnoaa package"
  ) +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE)
```





